### 2019-09-01 FakeNewsAnalysis/paperLatex/paperFelipeNevesDataScience.tex
initial paper commit

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{ Using News' Title and Content Text Features to Detect Fake News: a Viability Analysis}

\author{\IEEEauthorblockN{1\textsuperscript{st} Felipe dos Santos Neves}
\IEEEauthorblockA{\textit{DAINF} \\
\textit{UTFPR}\\
Curitiba, Brazil \\
fneves@alunos.utfpr.edu.br}
}

\maketitle

\begin{abstract}

The goal of this study was to define the viability of using features extracted from both titles 
and content from news article for the detection of Fake News using a large dataset and comparing
the results with these available on previous bibliography. 
\end{abstract}

\begin{IEEEkeywords}
Fake News, Classification, Data Science, Text Features, Unsupervised Learning
\end{IEEEkeywords}

\section{Introduction}

Automatic Fake News detection is a topic of growing interest in the last couple of years,
especially due to the influences presented in the 2016 American Elections \cite{allcott_gentzkow_2017},
however it is still very difficult to detect this type of content and stop its spread,
specially in social networks and messaging apps unregulated environments. These situations
alone give enough reason to pursue the definition of reliable approaches to Fake News detection.

In this paper we aim to utilize the results obtained from a previous study done specifically for Fake News 
about the 2016 American Election \cite{horne_2017} on a larger and more diverse dataset to determine
their viability in a more general scenario.

\section{Ease of Use}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
\centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

Test \cite{IEEEhowto:IEEEtranpage}
\bibliographystyle{ieeetr}
\bibliography{bibliography}

\end{document}

### 2019-09-03 FakeNewsAnalysis/paperLatex/paperFelipeNevesDataScience.tex
WIP paper

@@
 \usepackage{xcolor}
+\usepackage{multirow}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 
-\title{ Using News' Title and Content Text Features to Detect Fake News: a Viability Analysis}
+\title{ Using News' Title and Body Text Features to Detect Fake News: a Viability Study}
 
@@
 The goal of this study was to define the viability of using features extracted from both titles 
-and content from news article for the detection of Fake News using a large dataset and comparing
+and body from news article for the detection of Fake News using a large dataset and comparing
 the results with these available on previous bibliography. 
@@
 
-\section{Ease of Use}
+\section{Motivation}
 
-\subsection{Maintaining the Integrity of the Specifications}
+The main motivation of this work was to study possible implementations for the development
+of Fake News detection algorithms. Many previous studies have already proposed different
+algorithms for Fake News detection, but in some cases we found that there is the lack of a
+more extensive, reliable and unbiased dataset.
 
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
+A problem with using smaller datasets is that the results obtained over this dataset may not
+be similar to those obtained when using a bigger dataset because these smaller ones more often
+than not will present a limited range of variation in their subjects and contents, resulting
+in a more clustered distribution of parameters, making it easier for machine learning algorithms
+to achieve apparently good results.
 
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
+It was found that very interesting results were presented in Horne's paper \cite{horne_2017}, 
+however a very small dataset was used, and that is why we chose to use it's proposed features 
+to analyze a new and bigger dataset.
 
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
+\section{Previous Results}
 
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
-
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
+We use Horne's results as a baseline for our implementations. In his work he proposes the use of
+4 different features from both the title and the body of the news articles. These features are
+described in the Table \ref{table:horne_features}.
 
 \begin{table}[htbp]
-\caption{Table Type Styles}
+\caption{Features used in Horne's study}
 \begin{center}
-\begin{tabular}{|c|c|c|c|}
+\begin{tabular}{ |l|l|l| }
 \hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
+Section & \multicolumn{2}{ c| }{Feature Sets} \\
 \hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
+\multirow{4}{*}{Title} & LB & Lucas Radebe \\
+    & DC & Michael Duburry \\
+    & DC & Dominic Matteo \\
+    & RB & Didier Domi \\ \hline
+\multirow{3}{*}{Body} & MC & David Batty \\
+    & MC & Eirik Bakke \\
+    & MC & Jody Morris \\ \hline
 \end{tabular}
-\label{tab1}
+\label{table:horne_features}
 \end{center}
@@
 
-\begin{figure}[htbp]
-\centerline{\includegraphics{fig1.png}}
-\caption{Example of a figure caption.}
-\label{fig}
-\end{figure}
+In this paper he also implemented a SVM classifier to test the result of these features on his own
+proposed dataset. Considering that he only used 4 features, which is a considerably low number for
+a text classifier, he achieved a very impressive result with over 78\% accuracy in some tests , as
+shown in Table \ref{table:horne_results}.
 
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
+\begin{table}[htbp]
+\caption{Results achieved in Horne's study}
+\begin{center}
+\begin{tabular}{ |l|l|l| }
+\hline
+Feature Set & Baseline & Accuracy \\
+\hline
+Title & 50\% & 71\% \\
+Body & 50\% & 78\% \\
+\hline
+\end{tabular}
+\label{table:horne_results}
+\end{center}
+\end{table}
 
-\section*{Acknowledgment}
+For the purpose of comparing the results with a bigger dataset we chose to use a Fake News Corpus
+made available by M. Szpakowski \cite{szpakowski_2019} with over 8 million news articles, where around
+3 million of them were classified as Fake or Reliable. However due to limitations in time and 
+computational power available at the time of the development of this project we chose to pick
+only 750 entries of both Fake and Reliable news to reduce processing time, while still being able
+to achieve meaningful results.
 
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
+\section{Methods and Implementation}
 
-\section*{References}
+\section{Results}
 
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
+\section{Discussion}
 
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
+\section*{Acknowledgement}
+I would like to thank Prof. Luiz Celso Gomes Junior from UTFPR for the support
+provided during the development of this project.
 
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-\end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
-
-Test \cite{IEEEhowto:IEEEtranpage}
 \bibliographystyle{ieeetr}

### 2019-09-04 FakeNewsAnalysis/paperLatex/paperFelipeNevesDataScience.tex
Added plot script

@@
 \usepackage{multirow}
+\usepackage{hyperref}
+
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 \hline
-\multirow{4}{*}{Title} & LB & Lucas Radebe \\
-    & DC & Michael Duburry \\
-    & DC & Dominic Matteo \\
-    & RB & Didier Domi \\ \hline
-\multirow{3}{*}{Body} & MC & David Batty \\
-    & MC & Eirik Bakke \\
-    & MC & Jody Morris \\ \hline
+\multirow{4}{*}{Title} & per\_stop & Percentage of stop words\\
+    & NN & Number of nouns in the text \\
+    & avg\_wlen & Average word length \\
+    & FK & Flesch-Kincaid readability index \\ \hline
+\multirow{4}{*}{Body} & NN & Number of nouns in the text \\
+    & TTR & Token to type ration (lexical richness)\\
+    & WC & Number of words in the text \\
+    & quote & Number of quotes in the text \\ \hline
 \end{tabular}
@@
 
-In this paper he also implemented a SVM classifier to test the result of these features on his own
+In this paper he also implemented a \textit{SVM} classifier to test the result of these features on his own
 proposed dataset. Considering that he only used 4 features, which is a considerably low number for
@@
 \hline
-Title & 50\% & 71\% \\
-Body & 50\% & 78\% \\
+Title & 50\% & 78\% \\
+\hline
+Body & 50\% & 71\% \\
 \hline
@@
 
+For the development of this project we utilized the \textit{Python} language, with the aid of some external
+libraries as described in the Table \ref{table:python_libs}. \textit{Python} was chosen because it is widely
+considered the most accessible language for Data Science and it allows more people to be able to reproduce
+the results presented here \cite{cielen_meysman_mohamed_2016}\cite{richard_2018}. The full source
+code will also be made available at \url{https://github.com/Feupos/DataScience} for this purpose.
+This is a very important topic and will be further discussed in Section \ref{section:discussion}.
+
+\begin{table}[htbp]
+\caption{Python Libraries Utilized}
+\begin{center}
+\begin{tabular}{ |c|c| }
+\hline
+Library & Purpose \\ \hline
+Pandas & File processing and data structures \\ \hline
+scikit-learn & Machine learning algorithms \\ \hline
+NLTK & Natural language processing tools and dictionaries \\ \hline
+SentiStrength & Text polarity evaluation tool \\ \hline
+LexicalRichness & Lexical richness evaluation tool (TTR) \\ \hline
+Seaborn & Data visualization \\ \hline
+\end{tabular}
+\label{table:python_libs}
+\end{center}
+\end{table}
+
+With these tools we developed two scripts, one for extracting the features and another for testing
+machine learning algorithms for the classification of Fake News. Two types of classifiers were used:
+a linear kernel SVM (Space Vector Machine) classifier and a Random Forest classifier.
+
+We used \textit{scikit-learn}'s \textit{k-fold} cross validation for the performance evaluation of
+the classifiers.
+
+Besides the features already mentioned in \ref{table:horne_features} an additional set of features
+was used for the evaluating the performance of the algorithm, these features are listed on Table
+\ref{table:expanded_features} and will be referred as 'expanded feature set' whereas the features from
+Table \ref{table:horne_features} will be referred as 'original feature set'. In the same way the dataset
+from Horne's work \cite{horne_2017} will be referred as 'original dataset' and the dataset from Szpakowski
+\cite{szpakowski_2019} wil be referred as 'expanded dataset'.
+
+\begin{table}[htbp]
+\caption{Expanded Feature Set}
+\begin{center}
+\begin{tabular}{ |l|l|l| }
+\hline
+\multicolumn{2}{ |c| }{Features} \\
+\hline
+per\_stop & Percentage of stop words\\
+NN & Number of nouns in the text \\
+avg\_wlen & Average word length \\
+FK & Flesch-Kincaid readability index \\
+TTR & Token to type ration (lexical richness)\\
+WC & Number of words in the text \\
+quote & Number of quotes in the text \\
+NNP & Number of proper nouns in the text \\
+str\_neg & Average negative polarity of the text \\
+str\_pos & Average positive polarity of the text \\
+JJR & Number of comparative adjectives in the text \\
+JJS & Number of superlative adjectives in the text \\
+RBR & Number of comparative adverbs in the text \\
+RBS & Number of superlative adverbs in the text \\
+\hline
+\end{tabular}
+\label{table:expanded_features}
+\end{center}
+\end{table}
+
 \section{Results}
 
+During our tests we used multiple combinations of datasets and feature sets, to be able to highlight the
+difference in performance between them. We have also divided our results in 3 categories, one using only
+features extracted from the tiles, another with only features extracted from the bodies, and a third one
+combining all the extracted features from both titles and bodies.
+
+\subsection{Original Feature Set}
+
+In this test we used both the original feature set and the SVM classifier, to be able to reproduce the results
+found in Horne's work \cite{horne_2017}, however we were not able to achieve that, instead the results were
+'inverted', meaning that the accuracy using the Title features was achieved using the Body features and vice versa.
+This can be observed comparing our results from Table \ref{table:od_of} and the original result from Table \ref{table:horne_results}.
+
+We can also notice how the performance was significantly worse when using the expanded dataset, with the results
+on Table \ref{table:ed_of}.
+
+\begin{table}[htbp]
+\caption{Original Dataset and Original Features Performance}
+\begin{center}
+\begin{tabular}{ |l|l|l| }
+\hline
+Feature Set & Baseline & Accuracy \\
+\hline
+Title & 50\% & 72.67\% \\
+\hline
+Body & 50\% & 78.67\% \\
+\hline
+Combined & 50\% & 80\% \\
+\hline
+\end{tabular}
+\label{table:od_of}
+\end{center}
+\end{table}
+
+\begin{table}[htbp]
+\caption{Expanded Dataset and Original Features Performance}
+\begin{center}
+\begin{tabular}{ |l|l|l| }
+\hline
+Feature Set & Baseline & Accuracy \\
+\hline
+Title & 50\% & 58.13\% \\
+\hline
+Body & 50\% & 64.27\% \\
+\hline
+Combined & 50\% & 60.27\% \\
+\hline
+\end{tabular}
+\label{table:ed_of}
+\end{center}
+\end{table}
+
+\subsection{Expanded Feature Set}
+
+We also tried increasing the number of features in order to try to improve the performance,
+specially in the expanded dataset that had a very poor performance with the original features.
+In these tests we also switched the classifier for the Random Forest classifier due to the
+SVM being to slow to calculate with too many features and entries. Previous tests had already
+suggested that the classifier had little impact in the overall performance.
+The results are shown in the Table \ref{table:od_ef} and Table \ref{table:od_of}.
+
+\begin{table}[htbp]
+\caption{Original Dataset and Expanded Features Performance}
+\begin{center}
+\begin{tabular}{ |l|l|l| }
+\hline
+Feature Set & Baseline & Accuracy \\
+\hline
+Title & 50\% & 71.33\% \\
+\hline
+Body & 50\% & 82\% \\
+\hline
+Combined & 50\% & 82\% \\
+\hline
+\end{tabular}
+\label{table:od_ef}
+\end{center}
+\end{table}
+
+\begin{table}[htbp]
+\caption{Expanded Dataset and Expanded Features Performance}
+\begin{center}
+\begin{tabular}{ |l|l|l| }
+\hline
+Feature Set & Baseline & Accuracy \\
+\hline
+Title & 50\% & 61.93\% \\
+\hline
+Body & 50\% & 73.93\% \\
+\hline
+Combined & 50\% & 71.8\% \\
+\hline
+\end{tabular}
+\label{table:ed_ef}
+\end{center}
+\end{table}
+
+\subsection{Compared Results}
+
+TODO: plot new graph
+
 \section{Discussion}
+\label{section:discussion}
+
+\section{Discussion}
+\label{section:discussion}
+
+\section{Conclusion}
 

### 2019-09-04 FakeNewsAnalysis/paperLatex/plot.py

@@ -1,0 +1,16 @@
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+import pandas as pd
+
+
+data = [['Original Dataset', 'Original Features (Title)', 1],['Original Dataset', 'Expanded Features (Title)',2],
+        ['Original Dataset', 'Original Features (Body)', 1],['Original Dataset', 'Expanded Features (Title)',2],
+        ['Original Dataset', 'Original Features (Both)', 1],['Original Dataset', 'Expanded Features (Title)',2],
+        ['Expanded Dataset', 'Original Features (Title)', 3],['Expanded Dataset', 'Expanded Features (Title)',4],
+        ['Expanded Dataset', 'Original Features (Body)', 3],['Expanded Dataset', 'Expanded Features (Title)',4],
+        ['Expanded Dataset', 'Original Features (Both)', 3],['Expanded Dataset', 'Expanded Features (Title)',4]]
+df = pd.DataFrame(data,columns=['Dataset','Feature Set','Accuracy'],dtype=float)
+sns.barplot(x="Accuracy", y="Feature Set", hue="Dataset", data=df);
+
+plt.show()

### 2019-09-04 FakeNewsAnalysis/paperLatex/paperFelipeNevesDataScience.tex
improved plot and text

@@
 In these tests we also switched the classifier for the Random Forest classifier due to the
-SVM being to slow to calculate with too many features and entries. Previous tests had already
-suggested that the classifier had little impact in the overall performance.
+SVM being to slow to calculate with too many features and entries.
+
+Previous tests had already suggested that the classifier had little impact in the overall performance.
 The results are shown in the Table \ref{table:od_ef} and Table \ref{table:od_of}.
@@
 
-TODO: plot new graph
+For the purpose of being able to better distinguish the results on each step we show
+in Figure \ref{fig:results} a comparison between all the results side by side.
+
+\begin{figure}[htbp]
+\centerline{
+    \includegraphics[width=0.5\textwidth]
+    {content/results.png}
+    }
+\caption{Results comparison}
+\label{fig:results}
+\end{figure}
 
@@
 
-\section{Discussion}
-\label{section:discussion}
+As it is clear to see in Figure \ref{fig:results} it is clear that the performance over
+this bigger dataset was significantly worse than in the original paper \cite{horne_2017},
+however it is important to understand why whe had such difference.
+
+One very obvious reason for the divergence of the results of the classifier for the original
+dataset is that the algorithms, libraries and classifiers utilized in our implementation.
+This means that even though the author shared the dataset \cite{horne_data_2017}, as they
+did not share their implementation and it makes it very hard to reproduce the same results.
+
+In the field of Data Science reproducibility is a very important issue because due to the
+problem that finding reliable and unbiased datasets represent we often find the need to
+test and validate solutions proposed in other peoples' works. And in cases where either
+the source code or the datasets are not available it becomes near impossible to do
+\cite{peng_2011}\cite{peng_2015}.
+
+In this case the main problem with the original dataset  is that no only it had only 150
+entires, but also that it had only one subject, the 2016 American Elections, and what this
+means in terms of the features and classifiers is that the reliable news and fake news will
+share many similar characteristics among the items in their groups, an thus making it easier
+to achieve a better accuracy on the classifier output. It is easier to separate groups when
+their items are very similar inside each group, and share the same differences between items
+of the two groups.
+
+As we can see on Figure \ref{fig:titleNN}, some features did not have a  different distribution
+over the fake and reliable news, meaning that a classifier would not be able to separate these
+classes based on this feature only. Using more features helped overcome this problem but some
+features may actually decrease the overall performance.
+
+\begin{figure}[htbp]
+\centerline{
+    \includegraphics[width=0.5\textwidth]
+    {content/titleFK.png}
+    }
+\caption{FK index distribution over titles}
+\label{fig:titleNN}
+\end{figure}
+
+This also brings the problem that the features were chosen among the ones with the highest
+statistical significance in the original dataset, and this significance may not hold up
+in a different dataset such as ours. Another issue is that as the original dataset had the
+news manually chosen it is more susceptible to bias, and the fake news chosen were more
+explicit, with more distinguishable features compared to reliable news, than the ones in
+our expanded dataset.
 
@@
 
+The main point that we can drawn from this analysis is that the same feature set may not be
+efficient for classifying two different datasets. In this case the characteristics of the two
+datasets should be compared to see if they are similar enough. Therefore this approach of
+trying to use few but very specific features does not work well in the general case, because
+any change in the dataset may reduce the classification performance. To actually improve overall
+performance the best approach is to choose a different and wider set of features, and avoid choosing
+them based on a specific dataset, this way we can avoid the overfitting of the classification,
+where new inputs will have a worse performance due to being just slightly difference from the
+initial dataset.
+
+However only using more features is good but may have some drawbacks in regarding the processing
+speed that will be increased. Also, if the features set contains features that are not relevant
+to the analysis the performance may actually decrease due to containing inputs that cannot be
+separated by the classifier, such as what happened with the performance of out classification using
+the original and expanded features over the titles of the original dataset, where the performance
+dropped from 72.67\% with the original features to 71.33\% when using the expanded features, as shown
+on Table \ref{table:od_of} and Table \ref{table:od_ef} respectively.
+
+Another factor that can influence the performance is the choice of the classifier itself.
+The SVM is not appropriate for non-normalized data, because each input will influence the others
+and therefore the results may be good even though some features are having little impact in the
+classification. A more appropriate solution in this cases would be either normalize all the data
+or to utilize another type of classifier such as Random Forest, where the inputs do not influence
+each other and therefore requires no normalization. Also, when working with larger datasets the
+computational power required has a great impact, and the Random Forest algorithm resulted in
+significantly lower processing times, to the order of 100 times. But this of course cannot be
+generalized as the number of entries and configuration of each classifier can heavily impact
+on their speed.
+
 \section*{Acknowledgement}
+
 I would like to thank Prof. Luiz Celso Gomes Junior from UTFPR for the support
-provided during the development of this project.
+provided during the development of this project. It is also important to commend
+the work of all the people that collaborate developing free Data Science libraries
+and other resources that helps making Data Science more accessible, and without
+whom this work would not be possible.
 

### 2019-09-04 FakeNewsAnalysis/paperLatex/plot.py

@@ -3,14 +3,14 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 import pandas as pd
 
-
-data = [['Original Dataset', 'Original Features (Title)', 1],['Original Dataset', 'Expanded Features (Title)',2],
-        ['Original Dataset', 'Original Features (Body)', 1],['Original Dataset', 'Expanded Features (Title)',2],
-        ['Original Dataset', 'Original Features (Both)', 1],['Original Dataset', 'Expanded Features (Title)',2],
-        ['Expanded Dataset', 'Original Features (Title)', 3],['Expanded Dataset', 'Expanded Features (Title)',4],
-        ['Expanded Dataset', 'Original Features (Body)', 3],['Expanded Dataset', 'Expanded Features (Title)',4],
-        ['Expanded Dataset', 'Original Features (Both)', 3],['Expanded Dataset', 'Expanded Features (Title)',4]]
-df = pd.DataFrame(data,columns=['Dataset','Feature Set','Accuracy'],dtype=float)
-sns.barplot(x="Accuracy", y="Feature Set", hue="Dataset", data=df);
-
+data = [['Original', 'Original', 'Title', 72.67],['Original', 'Expanded', 'Title',71.33],
+        ['Original', 'Original', 'Body', 78.67],['Original', 'Expanded', 'Body',82],
+        ['Original', 'Original', 'Combined', 80],['Original', 'Expanded', 'Combined',82],
+        ['Expanded', 'Original', 'Title', 56.26],['Expanded', 'Expanded', 'Title', 61.93],
+        ['Expanded', 'Original', 'Body', 62.19],['Expanded', 'Expanded', 'Body', 73.93],
+        ['Expanded', 'Original', 'Combined', 61.48],['Expanded', 'Expanded', 'Combined', 71.8]]
+df = pd.DataFrame(data,columns=['Dataset','Features', 'Feature Set','Accuracy'])
+#sns.barplot(x='Accuracy', y='Dataset', hue='Feature Set', data=df, orient='v')
+g = sns.catplot(x="Features", y="Accuracy", hue="Feature Set", col="Dataset", data=df, kind="bar", height=4, aspect=.7, palette="Blues")
+g.set(ylim=(50, 85))
 plt.show()

### 2019-09-05 FakeNewsAnalysis/paperLatex/paperFelipeNevesDataScience.tex
finished paper

@@
 
-The goal of this study was to define the viability of using features extracted from both titles
-and body from news article for the detection of Fake News using a large dataset and comparing
-the results with these available on previous bibliography.
+In the past few years there have been a large number of algorithms developed to detect
+Fake News. We analyze the viability of using features extracted from both titles and
+body from news article for the detection of Fake News using a large dataset and comparing
+the results with these available on previous bibliography. We find that features used
+to classify a one dataset may not be appropriate to classify from a different one, due
+to the differences presented in these datasets. At last we test our model achieving a
+classification accuracy of 73.93\%.
+
 \end{abstract}
@@
 Automatic Fake News detection is a topic of growing interest in the last couple of years,
-especially due to the influences presented in the 2016 American Elections \cite{allcott_gentzkow_2017},
+especially due to the its impact in the 2016 American Elections \cite{allcott_gentzkow_2017}\cite{dey_rafi_parash_arko_chakrabarty_2018},
 however it is still very difficult to detect this type of content and stop its spread,
@@
 
-Besides the features already mentioned in \ref{table:horne_features} an additional set of features
+Besides the features already mentioned in Table \ref{table:horne_features} an additional set of features
 was used for the evaluating the performance of the algorithm, these features are listed on Table
@@
 this bigger dataset was significantly worse than in the original paper \cite{horne_2017},
-however it is important to understand why whe had such difference.
+however it is important to understand why we had such difference.
 
@@
 dropped from 72.67\% with the original features to 71.33\% when using the expanded features, as shown
-on Table \ref{table:od_of} and Table \ref{table:od_ef} respectively.
+on Table \ref{table:od_of} and Table \ref{table:od_ef} respectively. Simply adding more features
+will not necessarily cause an improvement in the performance \cite{volkova_shaffer_jang_hodas_2017}.
 
